
---
title: "Introduction to svatools"
<!-- output: rmarkdown::html_vignette -->
output: github_document
vignette: >
  %\VignetteIndexEntry{Introduction to svatools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
development:
  mode: auto
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# svatools

![GitHub License](https://img.shields.io/github/license/biopsichas/svatools)
![GitHub release (latest SemVer)](https://img.shields.io/github/v/release/biopsichas/svatools?display_name=tag&label=last%20release&logo=github&sort=semver)
![GitHub Release Date](https://img.shields.io/github/release-date/biopsichas/svatools?label=last%20release%20date&logo=github)

The goal of svatools is to help with [SWAT+ model](https://swat.tamu.edu/software/plus/) input data preparation, visualization and model output assessment. There are mostly functions, which were developed for the implementation of modeling tasks in the [OPTAIN project](https://www.optain.eu/). These tools are intended to fill the gaps in the SWAT+ workflow along side the main tools developed by [Christoph Schuerz](https://www.ufz.de/index.php?en=49467). Therefore, we highly recommend trying&using these tools:

- [SWATfarmR](http://chrisschuerz.github.io/SWATfarmR/) - R tool for preparing management schedules for SWAT model;
- [SWATplusR](https://chrisschuerz.github.io/SWATplusR/articles/SWATplusR.html) - R tool for sensitivity analyse, model calibration and validation;
- [SWATbuilder](https://git.ufz.de/users/sign_in) - R tool for building SWAT+ setups. 

## Installation

You can install the development version of svatools from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("biopsichas/svatools")
devtools::install_github("tkdweber/euptf2")
```

## Data

All the data required to run and test package is installed with package in extdata folder. Exact location on computer could be found running lines below. Please run it on your system to get it for you.   

```{r data}
library(svatools)
temp_path <- system.file("extdata", package = "svatools")
print(temp_path)
```

## Templates

In order to use *svatools* package functions with your data you should prepare your data to be inline with templates we have provided in *extdata* folder. Such are:

- **calibration_data.xlsx** - template for loading calibration (Q and WQ variables) data.
- **weather_data.xlsx** - template for loading weather variables.
- **soil_parameters.xlsx** - template for loading soil parameters dataset. 
- **GIS** - folder with GIS layers needed to run some functions.

Data prepared according to templates can be directly loaded into R and all the functions applied as described. 

## Example

Simple example to load and display data on one stations indented to be used in calibration. 

```{r example}
temp_path <- system.file("extdata", "calibration_data.xlsx", package = "svatools")
cal_data <- load_template(temp_path, 4326)
plot_fractions(cal_data$data, station = c("4"), c("PT"), c("P-PO4"))
```