---
title: "Weather data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading {#load_weather}

Loading weather data from templates could by done with the same [load_template](../reference/load_template.html) function as loading calibration data. However, data should for each meteorological station should be separated into different sheets. All data are loaded into nested list of list format, which could be used to run other functions. 

```{r met_load, message = FALSE, warning = FALSE}
library(SWATprepR)
temp_path <- system.file("extdata", "weather_data.xlsx", package = "SWATprepR")
met_lst <- load_template(temp_path, 3035)
```

## Plotting

Function [plot_weather](../reference/plot_weather.html) allows to plot meteorological data by selecting variables, time period and way of aggregating them. Time period should be selected one of *"second","minute","hour","day", "week", "month", or "year"*. While aggregation functions could be selected from `dplyr` package [summarise function](https://www.rdocumentation.org/packages/dplyr/versions/0.7.8/topics/summarise) available options.
All provided meteorological stations` data will be displayed. Relevant stations could be selected interactively on generated plot by selecting or deselecting stations in legend. *show all* or *hide all* buttons allows to select or deselect all stations. Press *Shift* + left mouse bottom and drag to zoom specific area, or press left mouse button and grad to explore plot, or press *Shift* + left and right mouse buttons at same time to zoom out. Toolbox in top right corner of figure could be used for these operations as well.  

```{r met_plot, message = FALSE, warning = FALSE}
plot_weather(met_lst, "PCP", "month", "sum")
```

Another function [plot_weather_compare](../reference/plot_weather_compare.html) allows comparing two meteorological datasets of same time scale. Potential examples of its usage could be different. One could be comparing different datasets in order to decide, which one should be used in modeling. Another example might be to examine how cleaning operations improved original data. Moreover, this function could be helpful in examining climate data downscaling or bias-correction results comparing them with observed data. Same options for aggregation as in [plot_weather_compare](../reference/plot_weather_compare.html) function can be used. 

```{r met_plot_comp, message = FALSE, warning = FALSE}
temp_path2 <- system.file("extdata", "weather_data_raw.xlsx", package = "SWATprepR")
met_lst2 <- load_template(temp_path2, 3035)
plot_weather_compare(met_lst, met_lst2, "PCP", "month", "mean", "clean", "raw")
```

## Interpolating

In cases where too few stations are available in a selected catchment or huge data gaps exists, interpolation could be applied between stations to cover those gaps. Example below loads basin boundary to be used in the modeling and adds with already [loaded](#load_weather) meteorological stations to the map. It could be seen that basin has just one station inside it. However, in this case station has only very few observations with huge gaps. Therefore, it is highly recommended in such cases collect data from nearby stations. These data can help to fill gaps and also could be used to generate spatial distribution for variables, which could be employed to propagate time series data for any number of virtual stations. Therefore, any number of meteorological stations could be prepared with full time series data. 

```{r met_station, message = FALSE, warning = FALSE}
library(mapview)
library(sf)
library(dplyr)
basin_path <- system.file("extdata", "GIS/basin.shp", package = "SWATprepR")
basin <-st_transform(st_read(basin_path), 4326) %>% 
  mutate(NAME = "Basin")
stations <- st_transform(met_lst$stations, 4326)
mapview(stations) + mapview(basin)
```

For such need function [interpolate](../reference/interpolate.html) was prepared. It allows to do interpolation between data in presented stations. Function's *grid_spacing* parameter defines, how dense spatial interpolation points will be. Interpolate function creates virtual stations. Data needed for this function are [loaded](#load_weather) stations observations, basin boundary shape path, DEM raster path. It is important, that GIS data used in function would be of same coordinate system.

Function parameters are *grid_spacing*, *p_vector* - character vector representing weather variables to interpolate (optional, default all variables selected c("PCP", "SLR", "RELHUM", "WNDSPD", "TMP_MAX", "TMP_MIN" ), *idw_exponent* - numeric value for exponent parameter to be used in interpolation  (optional, default value is 2). 

After running [interpolate](../reference/interpolate.html) function the [transform_to_list](../reference/transform_to_list.html) function could be used to transform interpolation results to list of list format, which is used with other functions (after loading weather data from template). So all the other functions can be applied. Transformation function requires dates for the beginning and the end of time series. This info can be extracted from loaded weather data (used in an interpolation) with the internal function `SWATprepR:::get_dates`. 

```{r met_interpolation, eval=FALSE, message = FALSE, warning = FALSE}
DEM_path <- system.file("extdata", "GIS/DEM.tif", package = "SWATprepR")
result <- SWATprepR::interpolate(met_lst, basin_path, DEM_path, 2000) 

##To transform interpolation results to list of lists, which could be used with other weather functions. 
start_date <- SWATprepR:::get_dates(met_lst)$min_date
end_date <- SWATprepR:::get_dates(met_lst)$max_date
met_lst_int <- transform_to_list(result, start_date, end_date)
```

### WGN input function

Weather generator inputs also can be directly prepared with [prepare_wgn](../reference/prepare_wgn.html) function from this package. 

```{r met_wgn_f, eval=FALSE, message = FALSE, warning = FALSE}
wgn <- prepare_wgn(met_lst, 
                   TMP_MAX = met_lst$data$ID10$TMP_MAX, 
                   TMP_MIN = met_lst$data$ID10$TMP_MIN, 
                   PCP = met_lst$data$ID9$PCP, 
                   RELHUM = met_lst$data$ID9$RELHUM, 
                   WNDSPD = met_lst$data$ID10$WNDSPD, 
                   MAXHHR = met_lst$data$ID11$MAXHHR, 
                   SLR = met_lst$data$ID9$SLR)
write.csv(wgn$wgn_st, "wgn_st.csv", row.names = FALSE, quote = FALSE)
write.csv(wgn$wgn_data, "wgn_data.csv", row.names = FALSE, quote = FALSE)
```

## Writing

### SWAT+ input and updating .sqlite database 

[add_weather](../reference/add_weather.html) function can be used to write all weather files into project's folder and update tables in project's sqlite database with weather, weather generator and relation to weather station information.  

```{r update, eval=FALSE, message = FALSE, warning = FALSE}
##Path to .sqlite
db_path <- "./output/test/project.sqlite"
add_weather(db_path, met_lst, wgn)
```
